---
title: "R Notebook"
output: html_notebook
---

Model with lag 1 of T10Y3M:
Model: real_gdp_growth = alpha + Beta1*T10Y3M_lag + epsilon

```{r}
reg_lag1 <- dynlm(real_gdp_growth~T10Y3M_lag, data = combined_data_filtered_ts)
summary(reg_lag1)
```

Conclusion: The once lagged values of T10Y3M are not significnat at the 5% level. We do not reject the null hypothesis that the value of T10Y3M cannot predict US real GDP growth 1 quarter ahead.

Model with lag 2 of T10Y3M:
Model: real_gdp_growth = alpha + Beta2*T10Y3M_lag2 + epsilon

```{r}
reg_lag2 <- dynlm(real_gdp_growth~T10Y3M_lag2, data = combined_data_filtered_ts)
summary(reg_lag2)
```

Conclusion: The twice lagged value of T10Y3M is significant for predicting real US GDP growth at the 5% level of significance. We reject the null hypothesis.


Model with lag 4 of T10Y3M:
Model: real_gdp_growth = alpha + Beta4*T10Y3M_lag4 + epsilon:

```{r}
reg_lag4 <- dynlm(real_gdp_growth~T10Y3M_lag4, data = combined_data_filtered_ts)
summary(reg_lag4)
```

Conclusion: The four times lagged values of T10Y3M are significant at the 5% level. We reject th

Unrestricted Model:
real_gdp_growth = alpha + beta1*T10Y3M_lag + beta2*T10Y3M_lag2 + beta3*T10Y3M_lag3 + beta4*T10Y3M_lag4 + epsilon, where epsilon is the eror term

Test for Stationarity - Augmented Dickey-Fuller Test for Real GDP Growth:
Model for ADF test:
delta_real_gdp_growth = alpha + beta1*real_gdp_growth_lag + beta2*delta_real_gdp_growth_lag + epsilon

```{r adf test gdp}
reg_gdp <- dynlm(delta_real_gdp_growth~real_gdp_growth_lag + delta_real_gdp_growth_lag, data = combined_data_filtered_ts)
summary(reg_gdp)
```

Conclusion:
The t value of real_gdp_growth_lag is -4.373, which is below the critical value of -2.9, so we reject the null hypothesis of non-stationarity of real GDP growth. Real GDP growth is stationary.

Test for Stationarity - Augmented Dickey-Fuller Test for T10Y3M:

```{r adf test T10Y3M}
reg_yield <- dynlm(delta_T10Y3M~T10Y3M_lag + delta_T10Y3M_lag, data = combined_data_filtered_ts)
summary(reg_yield)
```

Conclusion:
The t value of T10Y3M_lag is -3.216, which is below the critical value of -2.9, so we reject the null hypothesis of non-stationarity for T10Y3M. T10Y3M is stationary.

Walk-Forward Validation Attempts:
Does Work! Same results as when not using dynlm!
```{r test of expanding window walk-forward validation}
#test of expanding window walk-forward validation
set.seed(1)
wf_predictions <- rep(NA, nrow(combined_data_eval))

for (i in 1:nrow(combined_data_eval)) {
  #set training window
  train_window <- combined_data[1:(106 + i - 1), ]
  train_window_ts <- ts(train_window)
  #redo regression
  pred_reg <- dynlm(real_gdp_growth ~ T10Y3M_lag2 + real_gdp_growth_lag + 
                      real_gdp_growth_lag2, data = train_window_ts)
  #predictions
  wf_predictions[i] <- as.numeric(
    pred_reg$coefficients[1] + 
    pred_reg$coefficients[2]*combined_data[106 + i, 'T10Y3M_lag2'] +
    pred_reg$coefficients[3]*combined_data[106 + i, 'real_gdp_growth_lag'] +
    pred_reg$coefficients[4]*combined_data[106 + i, 'real_gdp_growth_lag2']
  )
}
```
```{r}
combined_data_ts <- ts(combined_data)
#test of expanding window walk-forward validation
#library("forecast")
set.seed(1)
wf_predictions <- rep(NA, nrow(combined_data_eval))
#wf_predictions <- vector(mode = "list", length = 10L)
#train_window1 <- ts(combined_data, end = combined_data$qtr_num[106]) #doesn't filter
#train_window1 <- ts(combined_data, end = 13970) #doesn't filter

#doesn't work
for (i in 1:nrow(combined_data_eval)) {
  train_window <- combined_data_ts[1:(106 + i - 1), ]
  #pred_window <- combined_data_ts[106 + i, ]
  pred_reg <- dynlm(real_gdp_growth ~ T10Y3M_lag2 + real_gdp_growth_lag + real_gdp_growth_lag2, data = train_window)
  
  #wf_predictions[i] <- forecast(pred_reg, h = 1) doesn't work
  #wf_predictions[i] <- predict(pred_reg, pred_window, n.ahead = 1) #doesn't work
  wf_predictions[i] <- pred_reg$coefficients[1] + pred_reg$coefficients[2]*combined_data_ts[106 + i, 'T10Y3M_lag2'] +
    pred_reg$coefficients[3]*combined_data_ts[106 + i, 'real_gdp_growth_lag'] +
    pred_reg$coefficients[4]*combined_data_ts[106 + i, 'real_gdp_growth_lag2']
}
````

```{r}
#make train_window a list and increment it
train_window <- vector(mode = "list", length = nrow(combined_data_eval)) #did work
#train_window[i] <- combined_data_ts[1:(106 + i - 1), ] doesn't work with ts
```







